### 2020-07-10 nb/preprocessing.Rmd
start making LeagueResults function

---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.4.2
  kernelspec:
    display_name: Python 3
    language: python
    name: python3
---

```{python}
import pandas as pd
import matplotlib.pyplot as plt
import json
import numpy as np

import warnings
warnings.filterwarnings('ignore')
```

```{python}
tag_df = pd.read_csv('../data/raw/tagcodes.csv')
tag_dict = dict(zip(tag_df.Tag, tag_df.Description))
```

```{python}
def load_json(file):
    "Load raw JSON data."
    with open(f'../data/raw/{file}.json','r') as f:
        return json.loads(f.read().replace('\\\\','\\').encode('utf-8'))
```

```{python}
def preprocess(national_league):
    "Create processed CSV from raw JSON for WyScout national league event-logs."

    # load JSON event-logs
    events = load_json(f'events_{national_league}')

    # create dataframe from JSON event-logs
    ev_df = pd.DataFrame(events)

    # decode tags
    for i in range(6):
        ev_df[f'tag{i}'] = ev_df['tags'].str[i].str['id'].map(tag_dict).str.lower()

    # decode player ids
    players = load_json('players')
    players_dict = dict(zip(pd.DataFrame(players).wyId,pd.DataFrame(players).shortName))
    ev_df['player'] = ev_df['playerId'].map(players_dict)

    # decode team ids
    teams = load_json('teams')
    teams_dict = dict(zip(pd.DataFrame(teams).wyId,pd.DataFrame(teams).name))
    ev_df['team'] = ev_df['teamId'].map(teams_dict)

    # unpack location and target
    ev_df['location'] = ev_df['positions'].str[0]
    ev_df['target'] = ev_df['positions'].str[1]

    # create simpler dataframe
    e_df = ev_df[['matchId','eventSec','matchPeriod','team','location','target','player','eventName','subEventName','tag0','tag1','tag2','tag3','tag4','tag5']]
    e_df['notes'] = e_df[['tag0','tag1','tag2','tag3','tag4','tag5']].values.tolist()
    e_df['tags'] = e_df['notes'].astype(str)
    e_df['accurate'] = e_df.tags.str.contains('accurate')
    e_df['accurate'] = np.where(e_df.tags.str.contains('not accurate'), 0, e_df.accurate)
    e_df.drop(columns=['tag0','tag1','tag2','tag3','tag4','tag5','notes'],inplace=True)

    # label in_play sequences
    play = 0
    plays = []
    next_action = []
    match_play = []
    matchId = e_df.iloc[0].matchId
    for i in range(len(e_df)):
        if i>1:
            row_minus2 = row_minus1
        if i>0:
            row_minus1 = row
        row = e_df.iloc[i]
        if i == 0:
            row_minus1 = row
            row_minus2 = row

        if row.eventName == 'Free Kick':
            # when the ball goes 'out of play', the game resumes with a 'free kick' (here includes Throw-Ins)
            play += 1
        elif 'goal' in row_minus2.tags and 'goal' in row_minus1.tags:
            # (there are generally two events marked goal -- the shot and the attempted save)
            play +=1

        if row.matchId == matchId:
            pass
        else:
            play = 0
            matchId = row.matchId
        plays.append(play)
        match_play.append(f'{matchId}-{play}')

    e_df['in_play'] = pd.Series(plays)
    e_df['match_play'] = pd.Series(match_play)

    # label possession sequences
    ps_sequence = np.full(len(e_df),'0000000000000000000000000000000000')
    played_to = np.full(len(e_df), '                                    ')
    to_team = np.full(len(e_df), '                                       ')

    for match in e_df.matchId.unique():
        match_df = e_df.loc[e_df.matchId == match]
        for team in match_df.team.unique():
            for play in match_df.in_play.unique():
                team_play_df = match_df.loc[match_df.in_play == play].loc[match_df.team == team]
                possession_sequence = 0
                for i in range(len(team_play_df)):
                    if i>0:
                        previous_row = row

                    row = team_play_df.iloc[i]

                    if i > 0:
                        if row.location != target:
                            possession_sequence += 1
                            played_to[previous_row.name] = 'Opposition' 
                            to_team[previous_row.name] = 'Opposition'
                        else:
                            played_to[previous_row.name] = row.player
                            to_team[previous_row.name] = row.team
                    team_name = team.split(' ')
                    team_abbr = ''
                    for word in team_name:
                        team_abbr = team_abbr + word[:3]
                    ps_sequence[row.name] = f'{match}-{play}-{team_abbr}-{possession_sequence}'
                    target = row.target

    e_df['possession'] = ps_sequence            
    e_df['played_to'] = played_to
    e_df['to_team'] = to_team

    e_df['played_to'] = np.where(e_df['played_to'] == '                                    ', 'Opposition', e_df['played_to'])
    e_df['to_team'] = np.where(e_df['to_team'] == '                                       ', 'Opposition', e_df['to_team'])

    # unpack x and y coorinates
    for foo in ('location','target'):
        for goo in ('x','y'):
            e_df[f'{foo}_{goo}'] = e_df[foo].str[goo]

    # change target coordinates of goal to inside goal
    e_df['target_x'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
                                100, e_df.target_x )
    e_df['target_y'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
                                50, e_df.target_y )

    e_df['subEventName'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
                                'Goal', e_df.subEventName )

    e_df['eventName'] = np.where(e_df['eventName'] == 'Others on the ball', 'On the Ball', e_df['eventName'])

    e_df['eventName'] = np.where(e_df['eventName'] == 'Duel', 'Challenge', e_df['eventName'])

    e_df['played_to'] = np.where(e_df.subEventName == 'Goal', 'Goal', e_df.played_to)
    e_df['to_team'] = np.where(e_df.subEventName == 'Goal', 'Goal', e_df.to_team)

    # change location coordinates of goal kick
    e_df['location_x'] = np.where((e_df.subEventName == 'Goal kick'), 5, e_df.location_x)
    e_df['location_y'] = np.where((e_df.subEventName == 'Goal kick'), 50, e_df.location_y)

    # add time in minutes and seconds
    e_df['minute'] = e_df.eventSec//60
    e_df.minute = e_df.minute.astype(int)
    e_df['seconds'] = e_df.eventSec % 60
    e_df.seconds = e_df.seconds.astype(int)
    e_df['time'] = e_df.minute.astype(str).str.zfill(2) + ':' + e_df.seconds.astype(str).str.zfill(2)
    e_df.drop(columns=['minute','seconds'],inplace=True)

    e_df.drop(columns=['location','target'], inplace=True)
    e_df.rename(columns={'player':'source','played_to':'target'},inplace=True)

    # include starting lineup data
    matches = load_json(f'matches_{national_league}')
    match_df = pd.DataFrame(matches)
    match_lineups = {}
    for match in match_df.wyId.unique():
        match_lineups[match] = {}
        for team in match_df.loc[match_df.wyId == match].teamsData.values[0].keys():
            match_lineups[match][teams_dict[int(team)]] = list(pd.Series(pd.DataFrame(match_df.loc[
                match_df.wyId == match].teamsData.str[str(team)].str['formation'].str['lineup'].values[0]).playerId.values).map(players_dict))

    def firstXI(row):
        return row.source in match_lineups[row.matchId][row.team]

    e_df['FirstXI'] = e_df.apply(firstXI, axis=1)

    # include home_or_away data
    h_a_map = {}
    for match in match_df.wyId.unique():
        h_a_map[match] = {}
        for team in match_df.loc[match_df.wyId == match].teamsData.values[0].keys():
            status = match_df.loc[match_df.wyId==match].teamsData.str[str(team)].str['side'].values[0].capitalize()
            h_a_map[match][team] = status

    def home_or_away(row):
        return h_a_map[row.matchId][str(row.teamId)]

    e_df['home_or_away'] = ev_df.apply(home_or_away, axis=1)
    e_df['to_team'] = np.where(e_df.to_team == e_df.team, e_df.home_or_away, e_df.to_team)
    opposition = {'Home':'Away','Away':'Home'}
    e_df['to_team'] = np.where(e_df.to_team == 'Opposition', e_df.home_or_away.map(opposition), e_df.to_team)
    e_df['to_team'] = np.where(e_df.tags.str.contains('own goal'), 'Own Goal', e_df.to_team)
    whose_goal = {'Home':'Home Goal', 'Away':'Away Goal'}
    e_df['to_team'] = np.where(e_df.to_team == 'Goal', e_df.home_or_away.map(whose_goal), e_df.to_team)
    whose_own_goal = {'Home':'Away Goal', 'Away':'Home Goal'}
    e_df['to_team'] = np.where(e_df.to_team == 'Own Goal', e_df.home_or_away.map(whose_own_goal), e_df.to_team)

    e_df.drop(columns=['tags','in_play'],inplace=True)

    e_df.to_csv(f'../data/processed/{national_league}FootballLeague2017-18.csv')
```

```{python}
national_leagues = ('England','France','Germany','Italy','Spain')
```

```{python}
for league in national_leagues:
    preprocess(league)
```

```{python}

```

```{python}
national_league='England'

# load JSON event-logs
events = load_json(f'events_{national_league}')

# create dataframe from JSON event-logs
ev_df = pd.DataFrame(events)

# decode tags
for i in range(6):
    ev_df[f'tag{i}'] = ev_df['tags'].str[i].str['id'].map(tag_dict).str.lower()

# decode player ids
players = load_json('players')
players_dict = dict(zip(pd.DataFrame(players).wyId,pd.DataFrame(players).shortName))
ev_df['player'] = ev_df['playerId'].map(players_dict)

# decode team ids
teams = load_json('teams')
teams_dict = dict(zip(pd.DataFrame(teams).wyId,pd.DataFrame(teams).name))
ev_df['team'] = ev_df['teamId'].map(teams_dict)

# unpack location and target
ev_df['location'] = ev_df['positions'].str[0]
ev_df['target'] = ev_df['positions'].str[1]

# create simpler dataframe
e_df = ev_df[['matchId','eventSec','matchPeriod','team','location','target','player','eventName','subEventName','tag0','tag1','tag2','tag3','tag4','tag5']]
e_df['notes'] = e_df[['tag0','tag1','tag2','tag3','tag4','tag5']].values.tolist()
e_df['tags'] = e_df['notes'].astype(str)
e_df['accurate'] = e_df.tags.str.contains('accurate')
e_df['accurate'] = np.where(e_df.tags.str.contains('not accurate'), 0, e_df.accurate)
e_df.drop(columns=['tag0','tag1','tag2','tag3','tag4','tag5','notes'],inplace=True)

# label in_play sequences
play = 0
plays = []
next_action = []
match_play = []
matchId = e_df.iloc[0].matchId
for i in range(len(e_df)):
    if i>1:
        row_minus2 = row_minus1
    if i>0:
        row_minus1 = row
    row = e_df.iloc[i]
    if i == 0:
        row_minus1 = row
        row_minus2 = row

    if row.eventName == 'Free Kick':
        # when the ball goes 'out of play', the game resumes with a 'free kick' (here includes Throw-Ins)
        play += 1
    elif 'goal' in row_minus2.tags and 'goal' in row_minus1.tags:
        # (there are generally two events marked goal -- the shot and the attempted save)
        play +=1

    if row.matchId == matchId:
        pass
    else:
        play = 0
        matchId = row.matchId
    plays.append(play)
    match_play.append(f'{matchId}-{play}')

e_df['in_play'] = pd.Series(plays)
e_df['match_play'] = pd.Series(match_play)

# label possession sequences
ps_sequence = np.full(len(e_df),'0000000000000000000000000000000000')
played_to = np.full(len(e_df), '                                    ')
to_team = np.full(len(e_df), '                                       ')

for match in e_df.matchId.unique():
    match_df = e_df.loc[e_df.matchId == match]
    for team in match_df.team.unique():
        for play in match_df.in_play.unique():
            team_play_df = match_df.loc[match_df.in_play == play].loc[match_df.team == team]
            possession_sequence = 0
            for i in range(len(team_play_df)):
                if i>0:
                    previous_row = row

                row = team_play_df.iloc[i]

                if i > 0:
                    if row.location != target:
                        possession_sequence += 1
                        played_to[previous_row.name] = 'Opposition' 
                        to_team[previous_row.name] = 'Opposition'
                    else:
                        played_to[previous_row.name] = row.player
                        to_team[previous_row.name] = row.team
                team_name = team.split(' ')
                team_abbr = ''
                for word in team_name:
                    team_abbr = team_abbr + word[:3]
                ps_sequence[row.name] = f'{match}-{play}-{team_abbr}-{possession_sequence}'
                target = row.target

e_df['possession'] = ps_sequence            
e_df['played_to'] = played_to
e_df['to_team'] = to_team

e_df['played_to'] = np.where(e_df['played_to'] == '                                    ', 'Opposition', e_df['played_to'])
e_df['to_team'] = np.where(e_df['to_team'] == '                                       ', 'Opposition', e_df['to_team'])

# unpack x and y coorinates
for foo in ('location','target'):
    for goo in ('x','y'):
        e_df[f'{foo}_{goo}'] = e_df[foo].str[goo]

# change target coordinates of goal to inside goal
e_df['target_x'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
                            100, e_df.target_x )
e_df['target_y'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
                            50, e_df.target_y )

e_df['subEventName'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
                            'Goal', e_df.subEventName )

e_df['eventName'] = np.where(e_df['eventName'] == 'Others on the ball', 'On the Ball', e_df['eventName'])

e_df['eventName'] = np.where(e_df['eventName'] == 'Duel', 'Challenge', e_df['eventName'])

e_df['played_to'] = np.where(e_df.subEventName == 'Goal', 'Goal', e_df.played_to)
e_df['to_team'] = np.where(e_df.subEventName == 'Goal', 'Goal', e_df.to_team)

# change location coordinates of goal kick
e_df['location_x'] = np.where((e_df.subEventName == 'Goal kick'), 5, e_df.location_x)
e_df['location_y'] = np.where((e_df.subEventName == 'Goal kick'), 50, e_df.location_y)

# add time in minutes and seconds
e_df['minute'] = e_df.eventSec//60
e_df.minute = e_df.minute.astype(int)
e_df['seconds'] = e_df.eventSec % 60
e_df.seconds = e_df.seconds.astype(int)
e_df['time'] = e_df.minute.astype(str).str.zfill(2) + ':' + e_df.seconds.astype(str).str.zfill(2)
e_df.drop(columns=['minute','seconds'],inplace=True)

e_df.drop(columns=['location','target'], inplace=True)
e_df.rename(columns={'player':'source','played_to':'target'},inplace=True)

# include starting lineup data
matches = load_json(f'matches_{national_league}')
match_df = pd.DataFrame(matches)
match_lineups = {}
for match in match_df.wyId.unique():
    match_lineups[match] = {}
    for team in match_df.loc[match_df.wyId == match].teamsData.values[0].keys():
        match_lineups[match][teams_dict[int(team)]] = list(pd.Series(pd.DataFrame(match_df.loc[
            match_df.wyId == match].teamsData.str[str(team)].str['formation'].str['lineup'].values[0]).playerId.values).map(players_dict))

def firstXI(row):
    return row.source in match_lineups[row.matchId][row.team]

e_df['FirstXI'] = e_df.apply(firstXI, axis=1)

# include home_or_away data
h_a_map = {}
for match in match_df.wyId.unique():
    h_a_map[match] = {}
    for team in match_df.loc[match_df.wyId == match].teamsData.values[0].keys():
        status = match_df.loc[match_df.wyId==match].teamsData.str[str(team)].str['side'].values[0].capitalize()
        h_a_map[match][team] = status

def home_or_away(row):
    return h_a_map[row.matchId][str(row.teamId)]

e_df['home_or_away'] = ev_df.apply(home_or_away, axis=1)
e_df['to_team'] = np.where(e_df.to_team == e_df.team, e_df.home_or_away, e_df.to_team)
opposition = {'Home':'Away','Away':'Home'}
e_df['to_team'] = np.where(e_df.to_team == 'Opposition', e_df.home_or_away.map(opposition), e_df.to_team)
e_df['to_team'] = np.where(e_df.tags.str.contains('own goal'), 'Own Goal', e_df.to_team)
whose_goal = {'Home':'Home Goal', 'Away':'Away Goal'}
e_df['to_team'] = np.where(e_df.to_team == 'Goal', e_df.home_or_away.map(whose_goal), e_df.to_team)
whose_own_goal = {'Home':'Away Goal', 'Away':'Home Goal'}
e_df['to_team'] = np.where(e_df.to_team == 'Own Goal', e_df.home_or_away.map(whose_own_goal), e_df.to_team)

e_df.drop(columns=['tags','in_play'],inplace=True)

```