### 2020-07-24 rmd/SocialNetworkAnalysis_R.Rmd
done almost

---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.4.2
  kernelspec:
    display_name: R
    language: R
    name: ir
---

```{r}
library(plyr) # data manipulation 
library(dplyr) # data manipulation
library(igraph) # for graph/network analysis 
library(ggraph) # graph visualization in the style of ggplot
library(netrankr) # network centrality
library(ggplot2) # for plotting visualizations
library(ggforce) # adds functionality to ggplot2, eg. geom_circle
library(ggrepel) # adds 'repellent' non-overlapping labels to ggplot
library(pheatmap) # prettier heat maps
library(RColorBrewer) # color palettes
library(docstring) # allows ?help to display function descriptions analogous to Pythonic docstrings
```

```{r}
LoadAndScaleData <- function(national_league) {
    #' Load preprocessed event logs for given national league.

    events <- read.csv(paste0('../data/processed/', national_league, '.csv'))

    x_scale <- 105
    y_scale <- 68

    events$location_x <- events$location_x/100 * x_scale
    events$target_x <- events$target_x/100 * x_scale

    events$location_y <- y_scale - (events$location_y/100 * y_scale)
    events$target_y <- y_scale - (events$target_y/100 * y_scale)

    return(events)
}

```

```{r}
LeagueResults <- function(league){
    #' Test data integrity by calculating results and table for given national league.

    teams <- sort(unique(league$team))

    league_table <- matrix(0, nrow=length(teams), ncol=9)
    rownames(league_table) <- teams
    colnames(league_table) <- c('Pos','Pld','W','D','L','GF','GA','GD','Pts')

    for (m in unique(league$matchId)){

        match <- league[league$matchId==m,]
        home_team <- unique(match[match$home_or_away=='Home',]$team)
        away_team <- unique(match[match$home_or_away=='Away',]$team)
        home_goals <- nrow(match[match$to_team=='Home Goal',])
        away_goals <- nrow(match[match$to_team=='Away Goal',])

        if (home_goals > away_goals){
            league_table[home_team,'W'] <- league_table[home_team,'W'] + 1
            league_table[away_team,'L'] <- league_table[away_team,'L'] + 1
        } else if (away_goals > home_goals){
            league_table[home_team,'L'] <- league_table[home_team,'L'] + 1
            league_table[away_team,'W'] <- league_table[away_team,'W'] + 1    
        } else if (home_goals == away_goals) {
            league_table[home_team,'D'] <- league_table[home_team,'D'] + 1
            league_table[away_team,'D'] <- league_table[away_team,'D'] + 1
        }

        league_table[home_team,'GF'] <- league_table[home_team,'GF'] + home_goals
        league_table[away_team,'GF'] <- league_table[away_team,'GF'] + away_goals
        league_table[home_team,'GA'] <- league_table[home_team,'GA'] + away_goals
        league_table[away_team,'GA'] <- league_table[away_team,'GA'] + home_goals

        for (team in c(home_team, away_team)){
            league_table[team,'GD'] <- league_table[team,'GF'] - league_table[team,'GA']
            league_table[team,'Pts'] <- 3*league_table[team,'W'] + 1*league_table[team,'D']
            league_table[team,'Pld'] <- league_table[team,'Pld'] + 1
        }

    }

    league_table <- league_table[order(-league_table[,'Pts']),]

    league_table[,'Pos'] <- 1:20

    return(league_table)
}
```

```{r}
DrawPitch <- function(lengthPitch=105, widthPitch=68, arrow=c("none", "r", "l"), theme=c("light", "dark", "grey", "grass")) {
    #' Draw regulation football pitch with penalty areas and centre circle.
    #' Adapted from https://github.com/JoGall/soccermatics

  # define colours by theme
  if(theme[1] == "grass") {
    fill1 <- "#008000"
    fill2 <- "#328422"
    colPitch <- "grey85"
    arrowCol <- "white"
    colText <- "white"
  } else if(theme[1] == "light") {
    fill1 <- "grey98"
    fill2 <- "grey98"
    colPitch <- "grey60"
    arrowCol = "black"
    colText <- "black"
  } else if(theme[1] %in% c("grey", "gray")) {    
    fill1 <- "#A3A1A3"
    fill2 <- "#A3A1A3"
    colPitch <- "white"
    arrowCol <- "white"
    colText <- "black"
  } else if(theme[1] == "dark") {  
    fill1 <- "#1C1F26"
    fill2 <- "#1C1F26"
    colPitch <- "white"
    arrowCol <- "white"
    colText <- "white"
  } else if(theme[1] == "blank") {
    fill1 <- "white"
    fill2 <- "white"
    colPitch <- "white"
    arrowCol <- "black"
    colText <- "black"
  }
  lwd <- 0.5

  border <- c(5, 5, 5, 5)

  # mowed grass lines
  lines <- (lengthPitch + border[2] + border[4]) / 13
  boxes <- data.frame(start = lines * 0:12 - border[4], end = lines * 1:13 - border[2])[seq(2, 12, 2),]

  # draw pitch
  p <- ggplot() +
    # background
    geom_rect(aes(xmin = -border[4], xmax = lengthPitch + border[2], ymin = -border[3], ymax = widthPitch + border[1]), fill = fill1) +
    # mowed pitch lines
    geom_rect(data = boxes, aes(xmin = start, xmax = end, ymin = -border[3], ymax = widthPitch + border[1]), fill = fill2) +
    # perimeter line
    geom_rect(aes(xmin = 0, xmax = lengthPitch, ymin = 0, ymax = widthPitch), fill = NA, col = colPitch, lwd = lwd) +
    # centre circle
    geom_circle(aes(x0 = lengthPitch/2, y0 = widthPitch/2, r = 9.15), col = colPitch, lwd = lwd) +
    # kick off spot
    geom_circle(aes(x0 = lengthPitch/2, y0 = widthPitch/2, r = 0.25), fill = colPitch, col = colPitch, lwd = lwd) +
    # halfway line
    geom_segment(aes(x = lengthPitch/2, y = 0, xend = lengthPitch/2, yend = widthPitch), col = colPitch, lwd = lwd) +
    # penalty arcs
    geom_arc(aes(x0= 11, y0 = widthPitch/2, r = 9.15, start = pi/2 + 0.9259284, end = pi/2 - 0.9259284), col = colPitch, lwd = lwd) +
    geom_arc(aes(x0 = lengthPitch - 11, y0 = widthPitch/2, r = 9.15, start = pi/2*3 - 0.9259284, end = pi/2*3 + 0.9259284), 
             col = colPitch, lwd = lwd) +
    # penalty areas
    geom_rect(aes(xmin = 0, xmax = 16.5, ymin = widthPitch/2 - 20.15, ymax = widthPitch/2 + 20.15), fill = NA, col = colPitch, lwd = lwd) +
    geom_rect(aes(xmin = lengthPitch - 16.5, xmax = lengthPitch, ymin = widthPitch/2 - 20.15, ymax = widthPitch/2 + 20.15), 
              fill = NA, col = colPitch, lwd = lwd) +
    # penalty spots
    geom_circle(aes(x0 = 11, y0 = widthPitch/2, r = 0.25), fill = colPitch, col = colPitch, lwd = lwd) +
    geom_circle(aes(x0 = lengthPitch - 11, y0 = widthPitch/2, r = 0.25), fill = colPitch, col = colPitch, lwd = lwd) +
    # six yard boxes
    geom_rect(aes(xmin = 0, xmax = 5.5, ymin = (widthPitch/2) - 9.16, ymax = (widthPitch/2) + 9.16), fill = NA, col = colPitch, lwd = lwd) +
    geom_rect(aes(xmin = lengthPitch - 5.5, xmax = lengthPitch, ymin = (widthPitch/2) - 9.16, ymax = (widthPitch/2) + 9.16), 
              fill = NA, col = colPitch, lwd = lwd) +
    # goals
    geom_rect(aes(xmin = -2, xmax = 0, ymin = (widthPitch/2) - 3.66, ymax = (widthPitch/2) + 3.66), fill = NA, col = colPitch, lwd = lwd) +
    geom_rect(aes(xmin = lengthPitch, xmax = lengthPitch + 2, ymin = (widthPitch/2) - 3.66, ymax = (widthPitch/2) + 3.66), 
              fill = NA, col = colPitch, lwd = lwd) +
    coord_fixed() +
    theme(rect = element_blank(), 
          line = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank())

  # add arrow
  if(arrow[1] == "r") {
    p <- p + 
      geom_segment(aes(x = 0, y = -2, xend = lengthPitch / 3, yend = -2), 
                   colour = arrowCol, size = 1.5, arrow = arrow(length = unit(0.2, "cm"), type="closed"), linejoin='mitre')
  } else if(arrow[1] == "l") {
    p <- p + 
      geom_segment(aes(x = lengthPitch, y = -2, xend = lengthPitch / 3 * 2, yend = -2), 
                   colour = arrowCol, size = 1.5, arrow = arrow(length = unit(0.2, "cm"), type="closed"), linejoin='mitre')
  }

  return(p)
}
```

```{r}
ShowMatchEvents <- function(events, match_id, team_name, home_or_away='Home', flip=F) {
    #' Visualize events on pitch for given match and team.

    # select match
    game <- events[events$matchId==match_id,]
    goals <- game[game$subEventName=='Goal',]

    # note teams
    teams <- unique(game$team)

    # limit events to team specified by name or home/away
    if(!missing(team_name)){
        game <- game[game$team==team_name,]
        home_or_away <- unique(game$home_or_away)
    } else {
        game <- game[game$home_or_away==home_or_away,]
        team_name <- unique(game$team)
    }

    # note opposition
    opposition <- teams[teams!=team_name]

    # get score
    team_score <- table(goals$team)[team_name]
    opposition_score <- table(goals$team)[opposition]

    # flip coordinates if desired
    if (flip==T) {
        game$location_x <- 105 - game$location_x
        game$location_y <- 68 - game$location_y
        direction_of_play = 'l'
    } else { 
        direction_of_play = 'r'
    }

    # limit attention to main events
    game_events <- game[game$eventName == 'Pass'| 
                        game$eventName == 'Shot' | 
                        game$eventName == 'On the Ball' | 
                        game$eventName == 'Challenge' | 
                        game$eventName == 'Free Kick',]

    # draw pitch
    p <- DrawPitch(theme='grey', arrow=direction_of_play) + 
        geom_point(data = game_events, 
               aes(location_x , location_y, fill=eventName, shape=eventName), pch=21, alpha=1, size=2 ) + 
        geom_point(data=game_events[game_events$subEventName=='Goal',], 
                   aes(location_x, location_y), shape=13, size=5) +
#         geom_label_repel(data=game_events[game_events$subEventName=='Goal',], 
#                          aes(location_x, location_y, label = paste0(source,'(',matchPeriod,' ',time,')')), label.padding=0.1, size=2.3, alpha=1) +
        theme(legend.direction='horizontal', legend.position=c(0.5,0)) +
        scale_fill_manual(values=c("red", "black", 'white', "yellow", 'blue'), name='Event Type') +
        ggtitle(paste('Event Map:', team_name,home_or_away,'versus',opposition,'(',team_score,'-',opposition_score,')'))

    p$figname <- paste0('EventMap',home_or_away)

    return(p)

}
```

```{r}
ShowPassesAndShots <- function(events, match_id, team_name, home_or_away='Home', flip=F) {
    #' Visualize passes on pitch for given match and team.

    # select match
    game <- events[events$matchId==match_id,]
    goals <- game[game$subEventName=='Goal',]

    # note teams
    teams <- unique(game$team)

    # limit events to team specified by name or home/away
    if(!missing(team_name)){
        game <- game[game$team==team_name,]
        home_or_away <- unique(game$home_or_away)
    } else {
        game <- game[game$home_or_away==home_or_away,]
        team_name <- unique(game$team)
    }

    # note opposition
    opposition <- teams[teams!=team_name]

    # get score
    team_score <- table(goals$team)[team_name]
    opposition_score <- table(goals$team)[opposition]

    # flip coordinates if desired
    if (flip==T) {
        game$location_x <- 105 - game$location_x
        game$target_x <- 105 - game$target_x
        game$location_y <- 68 - game$location_y
        game$target_y <- 68 - game$target_y
        direction_of_play = 'l'
    } else { 
        direction_of_play = 'r'
    }

    # limit attention to passes and shots
    passes <- game[game$eventName == 'Pass' & game$team == team_name,]
    shots <- game[(game$eventName=='Shot' | game$subEventName=='Goal') & game$team==team_name,]

    # draw pitch
    p <- DrawPitch(theme='grass', arrow=direction_of_play) + 
        geom_segment(data=na.exclude(passes), 
                     aes(x=location_x, y=location_y, xend=target_x, yend=target_y, color=subEventName), 
                     alpha=1, arrow = arrow(length = unit(0.1,"cm"))) +
#         geom_label_repel(data=game[game$subEventName=='Goal',], 
#                      aes(location_x, location_y, label = paste0(source,'(',matchPeriod,' ',time,')')), label.padding=0.1, size=2.3, alpha=1) + 
        geom_segment(data=na.exclude(shots), 
                 aes(x=location_x, y=location_y, xend=target_x, yend=target_y, color=subEventName), 
                 alpha=1, arrow = arrow(length = unit(0.1,"cm"))) +
        theme(legend.position=c(0.5,-0.01), legend.direction='horizontal',
              legend.background=element_rect(fill='#008000', linetype='solid')) +
        scale_color_manual(values=c("red", "black", 'blue', 'orange', "yellow", 'grey', 'purple', 'white','brown'), name='Pass Type') +
        ggtitle(paste('Pass Map:',team_name,home_or_away,'versus',opposition,'(',team_score,'-',opposition_score,')'))

    p$figname <- paste0('PassMap',home_or_away)

    return(p)
} 
```

```{r}
GoalSeqs <- function(events, match_id) {
    match <- events[events$matchId==match_id,]
    goals <- match[match$subEventName=='Goal',]
    goal_seqs <- unique(goals$possession)   
    return(goal_seqs)
}
```

```{r}
TabulateSequence <- function(events, possession_sequence) {
    # Return a markdown table starting from the row before a possession sequence begins, and continuing until it ends.

    start <- min(events[events$possession == possession_sequence,]$X)
    stop <- max(events[events$possession == possession_sequence,]$X) + 1
    table <- events[start:stop,]
    df <- data.frame(table$matchPeriod, table$time, table$team, table$source, table$subEventName)
    names(df) <- c('Half', 'Time', 'Team', 'Event Description', 'Player')

    return(df)
}
```

```{r}
SequenceOnPitch <- function(events, possession_sequence){
    #' Draw Possession Sequence on Pitch

    data <- events[events$possession == possession_sequence,]

    p <- DrawPitch(theme='grass') + 
        geom_label_repel(data = data, 
                         aes(location_x, location_y, label = paste(source,time)), 
                         label.padding=0.1, size=2.3, alpha=1) +
        geom_curve(data = data, 
                   aes(x = location_x, xend = target_x, 
                       y = location_y, yend = target_y*.99, col = X), 
                   show.legend=FALSE, size=1, alpha = 1) +
        geom_point(data = data, 
                   aes(location_x , location_y, shape=eventName), size=2) + 
        ggtitle(paste0(nrow(data),'-part ', unique(data$team)[1], 
                       ' Possession Sequence \nending with ', 
                       data[nrow(data),]$subEventName, ' by ', 
                       data[nrow(data),]$source))  +
        theme(legend.position=c(0.5,0.08), 
              legend.direction='horizontal', 
              legend.title=element_blank())

    return(p)
}
```

```{r}
SequenceGraph <- function(events, possession_sequence){
    #' Return Graph of Possession Sequence

    data <- events[events$possession == possession_sequence,]
    data <- data[data$source!='' & data$target!='nan',]
    nodes <- unique(c(as.character(data$source),as.character(data$target)))
    edges <- data.frame(data$source, data$target)
    g <- graph_from_data_frame(d=edges, vertices=nodes, directed=TRUE)
    g$id <- possession_sequence
    g$data <- data
    g$team <- as.character(unique(data$team))

    return(g)
}
```

```{r}
VisualizeGraph <- function(possession_sequence_graph) {
    #' Visualize Possession Sequence as Linear Graph with Looping Edges

    g <- possession_sequence_graph
    data <- possession_sequence_graph$data

    visualization <- ggraph(g, 'linear') + 
        geom_edge_arc(aes(color=data$eventName), 
                      arrow=arrow(length=unit(4,'mm')), 
                      fold=F,
                     width=1) +
        geom_edge_loop(aes(color=data$eventName),
                      width=1) +
        geom_node_point(color='black', 
                        size=2,
                        alpha=0.5) +
        geom_node_text(aes(label = name), 
                        repel=T, 
                        angle=90, hjust=2, ) +
        scale_edge_colour_manual(
            values=c('indianred3', 'wheat4', 'grey', 'grey30',
                    'red','blue','green','orange','purple','brown','pink'
                    ),
            name='Event Type') +
        theme_void()

    return(visualization)

}
```

```{r}
PassNetwork <- function(events, match_id, team_name, team_colour='red', home_or_away='Home', flip=F, lower_threshold=1, high_threshold=10) {
    #' Draw Pass Map of First XI with nodes placed on mean (x,y) pitch-coordinates.

    game <- events[events$matchId == match_id,]
    goals <- game[game$subEventName=='Goal',]

    # note teams
    teams <- unique(game$team)

    # limit events to team specified by name or home/away
    if(!missing(team_name)){
        game <- game[game$team==team_name,]
        home_or_away <- unique(game$home_or_away)
    } else {
        game <- game[game$home_or_away==home_or_away,]
        team_name <- unique(game$team)
    }

    # note opposition
    opposition <- teams[teams!=team_name]

    # get score
    team_score <- table(goals$team)[team_name]
    opposition_score <- table(goals$team)[opposition]

    firstXI <- game[game$FirstXI == 'True',]
    mean_positions <- firstXI[firstXI$location_x>0 & firstXI$location_y>0 & firstXI$location_x<105 & firstXI$location_y<68,] %>% 
        group_by(team, matchId, source) %>%
          dplyr::summarise(x_mean = mean(location_x), y_mean = mean(location_y)) %>% 
          ungroup() %>%
          mutate(team = as.factor(team), id = as.factor(matchId)) %>%
          as.data.frame()

    pass_counts <- ddply(data.frame(game$source, game$target),.(game.source,game.target),nrow)
    names(pass_counts) <- c('source','target','passcount')
    step1 <- merge(mean_positions, pass_counts, by='source')
    step2 <- step1[,c(1,4,5,7,8)]
    names(step2)[2:3] <- c('source_x','source_y')
    names(mean_positions)[3] <- 'target'
    step3 <- merge(mean_positions, step2, by='target')
    team <- step3[step3$team==team_name,]

    if (flip==T) {
        team$source_x <- 105 - team$source_x
        team$x_mean <- 105 - team$x_mean
        team$source_y <- 68 - team$source_y
        team$y_mean <- 68 - team$y_mean
        mean_positions$x_mean <- 105 - mean_positions$x_mean
        mean_positions$y_mean <- 68 - mean_positions$y_mean
    }

    p <- (DrawPitch() +
        geom_segment(data=team[team$passcount>=lower_threshold,], 
                     size=1, colour=team_colour,
                     aes(x=source_x, y=source_y, 
                         xend=x_mean, yend=y_mean, alpha=passcount)) +
        geom_segment(data=team[team$passcount>=high_threshold,], 
                     size=1.5, colour='black',
                     aes(x=source_x, y=source_y, 
                         xend=x_mean, yend=y_mean, alpha=passcount)) +
        geom_label_repel(data = mean_positions[mean_positions$team==team_name,], 
                         aes(x_mean, y_mean, label = target), 
                         label.padding=0.5, size=4, alpha=0.8) +
        geom_point(data=team, aes(x_mean, y_mean,), 
                   fill=team_colour, colour='black', pch=21, size=3) +
        ggtitle(paste('Passing Network:',team_name,home_or_away,
                      'versus',opposition,'(',team_score,'-',opposition_score,')')) +
        theme(legend.position=c(0.5,0.07), legend.direction='horizontal'))

    p$figname <- paste0('PassingNetwork',home_or_away)

    return(p)
}
```

```{r}
GameGraph <- function(events, match_id, team_name, home_or_away='Home') {
    #' Return Possession Graph for given Match and Team.

    game <- events[events$matchId == match_id,]
    game <- game[game$source!='' & game$target!='' & game$target!='nan',]

     # note teams
    teams <- unique(game$team)

    # limit events to team specified by name or home/away
    if(!missing(team_name)){
        team_game <- game[game$team==team_name,]
        home_or_away <- unique(team_game$home_or_away)
    } else {
        team_game <- game[game$home_or_away==home_or_away,]
        team_name <- as.character(unique(team_game$team))
    }

    opposition_game <- game[game$team != team_name,]
    opposition_team <- as.character(unique(opposition_game$team))

    team_nodes <- unique(c(as.character(team_game$source),as.character(team_game$target)))
    team_edges <- data.frame(team_game$source, team_game$target)
    team_graph <- graph_from_data_frame(d=team_edges, vertices=team_nodes, directed=TRUE)
    team_adj <- as.matrix(team_graph[])

    possession_seqs <- unique(team_game$possession)

    seq_start <- c()

    for (pseq in possession_seqs){
        if (sum(team_game$possession==pseq)>1){
            player <- as.character(team_game[team_game$possession==pseq,][1,]$source)
            if (player!='' & player!='nan'){
                seq_start <- append(seq_start, player)
            }
        }
    }

    start_counts <- table(as.factor(seq_start))

    for (i in (2:(length(start_counts)-1))){
        team_adj[nrow(team_adj)-1,][names(start_counts[i])] <- as.numeric(start_counts[i])
    }

    opposition_status <- as.character(unique(opposition_game$home_or_away))
    opp_to_opp <- as.numeric(summary(game[game$team != team_name,]$to_team)[opposition_status])

    opp_nodes <- unique(c(as.character(opposition_game$source),as.character(opposition_game$target)))
    opp_edges <- data.frame(opposition_game$source, opposition_game$target)
    opp_graph <- graph_from_data_frame(d=opp_edges, vertices=opp_nodes, directed=TRUE)
    opp_adj <- as.matrix(opp_graph[])

    opp_to_opp <- sum(opp_adj[,ncol(opp_adj)-1])
    team_adj[(nrow(team_adj)-1),(ncol(team_adj)-1)] <- opp_to_opp

    team_status <- unique(team_game$home_or_away)
    opposition_own_goals <- as.numeric(nrow(game[game$home_or_away==opposition_status & game$to_team==paste(team_status,'Goal'),]))
    team_adj[(nrow(team_adj)-1),(ncol(team_adj))] <- opposition_own_goals

    team_graph <- graph_from_adjacency_matrix(team_adj, mode='directed')

    team_graph$team <- team_name
    team_graph$opposition <- opposition_team
    team_graph$status <- home_or_away

    return(team_graph)
}
```

```{r}
AdjacencyMatrix <- function(graph){
    #' Return Adjacency Matrix for Graph

    return(as.matrix(graph[]))
}
```

```{r}
TransitionMatrix <- function(graph){
    #' Return Transition Matrix for Graph considered as time-homogeneous Markov Process

    matrix <- as.matrix(graph[])

    for (i in 1:nrow(matrix)){
        matrix[i,] <- matrix[i,]/sum(matrix[i,])
    }

    matrix[nrow(matrix),ncol(matrix)] <- 1
    matrix[nrow(matrix),] <- rep(0, ncol(matrix))

    # treat 'Goal' as absorbtion state
    matrix[nrow(matrix),ncol(matrix)] <- 1

    return(matrix)
}
```

```{r}
MatrixHeatMap <- function(matrix, color, number_format, title){
    #' Show Matrix as HeatMap

    hm <- pheatmap(matrix,color=color,
                   cluster_rows=F,cluster_cols=F,legend=F,
                    display_numbers=T,number_format=number_format,
                   fontsize_number=9,angle_col='315',
                    main=title)

    return(hm)
}
```

```{r}
ExpectedScoringTime <- function(transition_matrix){
    #' Calculate expected scoring time based on transition matrix.

    max_steps <- 10000
    hitting_times <- rep(0,(ncol(transition_matrix)-2))

    for (i in (1:(ncol(transition_matrix)-2))){

        state_probabilities <- matrix(NA,
                                      nrow=max_steps+1,
                                      ncol=ncol(transition_matrix),
                        dimnames=list(0:max_steps,(ncol(transition_matrix)-1):0))
        vector <- rep(0,ncol(transition_matrix))

        vector[i] <- 1
        state_probabilities[1,] <- vector

        for ( kk in 1:max_steps ) {
            state_probabilities[kk+1,] <- t(transition_matrix)%*%state_probabilities[kk,]
        }

        probs <- diff(state_probabilities[,ncol(transition_matrix)])
        hitting_time <- sum(probs*seq_along(probs))

        hitting_times[i] <- hitting_time
    }

    names(hitting_times) <- names(transition_matrix[1,])[1:(nrow(transition_matrix)-2)]

    return(hitting_times)

}
```

```{r}
SeasonAdjacencyMatrix <- function(events, team) {
    #' Return Adjacency Matrix for Team's full season.

    players <- unique(events[events$team==team & events$source!='' & events$target!='',]$source)
    n <- length(players) + 2
    squad_adj <- matrix(0,nrow=n,ncol=n)
    rownames(squad_adj) <- c(as.character(players), 'Opposition', 'Goal')
    colnames(squad_adj) <- rownames(squad_adj)

    games <- unique(events[events$team==team,]$matchId)

    for (game in games){

        # get graph and adjacency matrix for game
        game_graph <- GameGraph(events, game, team)
        game_adj <- AdjacencyMatrix(game_graph)

        # add values of game adjacency matrix to full-season matrix
        for (source in rownames(game_adj)){
            for (target in colnames(game_adj)){
                squad_adj[source,target] <- squad_adj[source,target] + game_adj[source,target]
            }
        }
    }

    return(squad_adj)
}
```

```{r}
MarkovExpectedGoals <- function(events, season_adjacency_matrix) {
    #' Calculate MXG for each player from team's Season Adjacency Matrix.

    season_graph <- graph_from_adjacency_matrix(season_adjacency_matrix)
    season_transition <- TransitionMatrix(season_graph)
    seasonXST <- ExpectedScoringTime(season_transition)
    seasonMXG <- 1/seasonXST
    stdize <- seasonMXG/(mean(seasonMXG))
    avg_events_per_game <- sum(season_adjacency_matrix)/38
    scale <- stdize * avg_events_per_game - avg_events_per_game

    return(scale)
}
```

```{r}
MXGTable <- function(events, team){
    #' Return dataframe with Degree, Loops, Goals, and MXG for each player in team.

    Team <- team
    m <- SeasonAdjacencyMatrix(events,Team)
    Degree <- rowSums(m)[1:(nrow(m)-2)]
    Loops <- diag(m)[1:(nrow(m)-2)]
    Goals <- m[,ncol(m)][1:(nrow(m)-2)]
    Player <- rownames(m)[1:(nrow(m)-2)]
    MXG <- MarkovExpectedGoals(events, m)[1:(nrow(m)-2)]
    df <- data.frame(Player, Team, MXG, Goals, Loops, Degree)   
    rownames(df) <- c()

    return(df)
}

```

```{r}
CalculateMXG <- function() {
    #' Get MXG and other stats for each player in all leagues.

    mxgtable_df <- data.frame(Player = character(),
                    Team = character(),
                    MXG = double(),
                    Goals = integer(),
                    Loops = integer(),
                    Degree = integer())

    leagues = c('England','France','Germany','Italy','Spain')
    for (league in leagues){
        events <- LoadAndScaleData(league)
        for (team in unique(events$team)){
            mxgtable_df <- rbind(mxgtable_df, MXGTable(events,team))

        }
    }

    write.csv(mxgtable_df[order(-mxgtable_df[,'MXG']),], file='../tables/MXG.csv', row.names=F)

}
```

```{r}
TallyEventTypes <- function(events, eventType=F){
    #' Return dataframe with tallies for specified type.

    if (eventType!=F){
        events <- events[events$eventName==eventType,]
        df <- data.frame(table(events$subEventName))
        df <- df[df$Freq>0,]
    } else {
        df <- data.frame(table(events$eventName))
    }

    df <- df[order(-df$Freq),]
    colnames(df)[1:2] <- c('Description','Frequency')

    df$Description <- as.character(df$Description)

    total <- sum(df$Frequency)
    df[nrow(df)+1,] <- c('**Total**',total)
    df$Frequency <- as.numeric(df$Frequency)

    df$Pct <- df$Frequency / total * 100
    rownames(df) <- c()
    colnames(df)[3] <- '%'

    df[,3] <- formatC(signif(df[,3],digits=3), digits=3,format="fg", flag="#")

    df[,'Frequency (%)'] <- paste0(as.character(
                            prettyNum(df$Frequency, big.mark=",", scientific=F)),
                           ' (', as.character(df[,3]),'%)')

    df <- df[,c(1,4)]

    return(df)
}
```

```{r}
ExploratoryTallies <- function(){
    #' Get exploratory tallies.

    events <- LoadAndScaleData('England')

    leagues = c('England','France','Germany','Italy','Spain')
    for (league in leagues[2:5]){
        events <- rbind(events, LoadAndScaleData(league))
    }

    types <- TallyEventTypes(events)
    passes <- TallyEventTypes(events,'Pass')
    shots <- TallyEventTypes(events,'Shot')

    write.csv(types,'../tables/EventTypeTally.csv',row.names=F)
    write.csv(passes,'../tables/PassTally.csv',row.names=F)
    write.csv(shots,'../tables/ShotTally.csv',row.names=F)

}
```

```{r}
SavePairedFigures <- function(league, match_id){
    #' Generate and Save Paired Figures.

    events <- LoadAndScaleData(league)

    EventMapHome <- ShowMatchEvents(events, match_id)
    EventMapAway <- ShowMatchEvents(events, match_id, home_or_away='Away', flip=T)

    PassMapHome <- ShowPassesAndShots(events, match_id)
    PassMapAway <- ShowPassesAndShots(events, match_id, home_or_away='Away', flip=T)

    PassingNetworkHome <- PassNetwork(events, match_id, team_colour='blue')
    PassingNetworkAway <- PassNetwork(events, match_id, home_or_away='Away', 
                                      team_colour='red',flip=T)

    pairedFigures <- list(EventMapHome, EventMapAway,
                      PassMapHome, PassMapAway,
                      PassingNetworkHome, PassingNetworkAway)

    for (fig in pairedFigures){
        png(filename=paste0('../figures/',fig$figname,'.png'),
            width=800, 
            height=650,
            units='px',
           pointsize=4,
           res=140)
        print(fig)
        dev.off()

    }   
}

```

```{r}
CountSequences <- function(events) {
    #' Count Possession Sequences and return DataFrame.

    poss_seqs <- data.frame(table(events$possession))
    df_psq <- data.frame(table(poss_seqs$Freq))
    goals_seqs <- events[events$subEventName == 'Goal',]$possession
    goals_seqs_events <- events[events$possession %in% goals_seqs,]
    df_gpsq0 <- data.frame(table(goals_seqs_events$possession))
    df_gpsq <- df_gpsq0[df_gpsq$Freq>0,]
    df_g <- data.frame(table(df_gpsq$Freq))

    m <- rbind(1:10, df_psq$Freq[1:10], df_g$Freq[1:10])
    mdf <- data.frame(m)
    rownames(mdf) <- c('Sequence Length', 'Frequency', 'Goal Scoring')

    total_goals <- sum(df_g$Freq)
    total_seqs <- sum(df_psq$Freq)
    final <- c('**Total**', total_seqs, total_goals)
    mdf <- cbind(mdf,final)
    colnames(mdf) <- c()

    df <- t(mdf)
    df[,2] <- prettyNum(df[,2], big.mark=",", scientific=F)
    df[,3] <- prettyNum(df[,3], big.mark=",", scientific=F)
    return(df)
}
```

```{r}
PossessionSequence <- function(league, sequence){
    #' Generate Figures and Table for given Possession Sequence.

    events <- LoadAndScaleData(league)
    write.csv(TabulateSequence(events, sequence),'../tables/PossessionSequence.csv',row.names=F)
    png(filename=paste0('../figures/PossessionSequence.png'),
        width=800, 
        height=600,
        units='px',
       pointsize=12,
       res=140)
    print(SequenceOnPitch(events, sequence))
    dev.off()

    g <- SequenceGraph(events, sequence)
    png(filename=paste0('../figures/PossessionGraph.png'),
        width=800, 
        height=600,
        units='px',
       pointsize=12,
       res=140)
    print(VisualizeGraph(g))
    dev.off()
}
```

```{r}
Top20MXG <- function(){
    #' Save CSV with top 20 players by MXG score.

    d <- read.csv('../tables/MXG.csv')
    top <- d[d$Degree>1000,]
    rownames(top) <- c()
    top['Loops/Degree'] <- top$Loops/top$Degree
    top['Goals/Degree'] <- top$Goals/(top$Degree)
    t <- top[c('Player','Team','Degree','Goals',
               'Goals/Degree','Loops','Loops/Degree','MXG')]
    write.csv(t[1:20,], '../tables/2.csv',row.names=F)
}

```

```{r}
SaveMatrixHeatmaps <- function(league, match_id, team_name){
    #' Save Heatmaps for Adjacency and Transition Matrices.

    events <- LoadAndScaleData(league)
    g <- GameGraph(events, match_id, team_name)
    adj_m <- AdjacencyMatrix(g)
    tr_m <- TransitionMatrix(g)

    matrix_cols <- c('darkgrey',colorRampPalette(c('white','red'))(35))
    adj_m_hm <- MatrixHeatMap(adj_m[1:11,1:11], color=matrix_cols, number_format='%.0f',
                  title=paste('Adjacency Matrix for',g$team,'(Starting XI)\n',
                              g$status,'vs.',g$opposition))
    adj_m_hm$type <- 'Adjacency'

    tr_m_hm <- MatrixHeatMap(tr_m, color=matrix_cols[2:36], number_format='%.2f',
                  title=paste('Transition Matrix for',g$team,g$status,'vs.',g$opposition))
    tr_m_hm$type <- 'Transition'

    hm_vec <- list(adj_m_hm, tr_m_hm)

    for (hm in hm_vec){
        png(filename=paste0('../figures/',hm$type,'MatrixAway.png'),
            width=1600, 
            height=1600,
            units='px',
           pointsize=12,
           res=240)

        print(hm)

        dev.off()   
    }

}
```

```{r}
GenerateTablesAndFigures <- function(league, match_id, possession_seq, team){
    #' Generate Tables and Figures for Paper.

    ExploratoryTallies()
    SavePairedFigures(league, match_id)
    PossessionSequence(league, possession_seq)
    SaveMatrixHeatmaps(league, match_id, team)
    CalculateMXG()
    Top20MXG()
}
```

```{r}
GenerateTablesAndFigures('England','2500032','2500032-91-Liv-0','Liverpool')
```